---
title: "Nuclear weapons"
author: "James Goldie, 360info"
---

```{r setup}
library(tidyverse)
library(glue)
library(sf)
library(ggsflabel)
library(gganimate)
library(themes360info)
library(rnaturalearth)
library(here)
```

```{r import}
stockpiles <- read_csv(here("data", "nuclear-warhead-stockpiles.csv"))
stockpiles
```

The OWID data is already really tidy - no need to do much with the country names, and ISO3 codes are already provided. (There are also relatively few nuclear powers.)

The current nuclear powers are `r stockpiles %>% filter(Year == max(Year, na.rm = TRUE)) %>% pull(Entity) %>% unique() %>% glue_collapse(sep = ", ", last = " and ")`.

Let's get our world map boundaries:

```{r boundaries}
sf::sf_use_s2(FALSE)

ne_countries(returnclass = "sf") %>%
# geoboundaries() %>%
  select(Code = iso_a3) -># %>%
  # st_make_valid() ->
boundaries

sf::sf_use_s2(TRUE)

# get the centroids of the nuclear powers
boundaries %>%
  filter(Code %in% unique(stockpiles$Code)) %>%
  st_make_valid() %>%
  st_centroid(of_largest_polygon = TRUE) ->
centroids

# now simplify the geometry for plotting
# boundaries %>%
#   st_simplify(dTolerance = 1) ->
# simple_boundaries
```

Finally, we need to join the nuclear data with the centroids:

```{r joinnuclear}
centroids %>%
  right_join(stockpiles, by = "Code") %>%
  filter(nuclear_weapons_stockpile > 0) ->
stockpile_points

# finally, let's split based on large and small stockpiles so that we can tweak
# the label appearance
```

Okay, let's plot:

```{r map}

si_count <- scales::label_number_si()
comma_count <- scales::label_comma(accuracy = 1)

warheads_map <- function(df, year) {
  warhead_plot <- ggplot(df) +
    # world map boundaries
    geom_sf(data = boundaries, fill = "#cccccc", colour = NA) +
    # points and labels
    geom_sf(aes(size = Count, alpha = Count), colour = "orange") +
    # labels: uk and france overlaps are handled manually to avoid jumping around (this would be way easier if gganimate was working)
    geom_sf_text(
      aes(label = pt_label),
      data = df %>% filter(Entity %in% c("United Kingdom")),
      family = "Body 360info", fontface = "bold",
      size = 3, direction = "both",
      # check_overlap = TRUE,
      hjust = "right", nudge_x = -2.5,
      colour = "black", show.legend = FALSE) +
    geom_sf_text(
      aes(label = pt_label),
      data = df %>% filter(Entity %in% c("France")),
      family = "Body 360info", fontface = "bold",
      size = 3, direction = "both",
      hjust = "left", nudge_x = 2.5,
      colour = "black", show.legend = FALSE) +
    # everyone else
    geom_sf_text(
      aes(label = pt_label),
      data = df %>% filter(!(Entity %in% c("United Kingdom", "France"))),
      family = "Body 360info", fontface = "bold",
      size = 3,
      # check_overlap = TRUE,
      # hjust = 0, nudge_x = 0.5,
      colour = "black", show.legend = FALSE) +
    # geom_text_repel(
    #   aes(x, y, label = pt_label),
    #   family = "Body 360info", fontface = "bold",
    #   size = 3, show.legend = FALSE, max.overlaps = 100,
    # ) +
    scale_alpha(
      limits =
        c(1, max(stockpile_points$nuclear_weapons_stockpile, na.rm = TRUE)),
      range = c(1, 0.35), guide = NULL) +
    scale_radius(name = "Num. warheads",
      limits = c(1,
        max(stockpile_points$nuclear_weapons_stockpile, na.rm = TRUE)),
      range = c(1, 80), labels = si_count, guide = NULL) +
    coord_sf(crs = "+proj=eck4", clip = "off") +
    theme_360() +
    theme(
      # panel.background = "white",
      plot.background = element_rect(fill = "white"),
      # panel.grid = element_blank(),
      legend.direction = "horizontal",
      legend.position = "top"
    ) +
    labs(
      x = NULL, y = NULL,
      title = paste("WARHEADS IN", year),
      subtitle = "A SUBTITLE HERE",
      caption = paste(
        "**CHART:** James Goldie, 360info",
        "**DATA:** Our World In Data [ourworldindata.org/nuclear-weapons]",
        sep = "<br>"
      )
    )

  save_360plot(warhead_plot, here("out", paste0("warheads-", year, ".png")),
    shape = "sdtv-landscape")
  return(warhead_plot)
}

# get the geometry coordinates
coordinates <- st_coordinates(stockpile_points)

stockpile_points %>%
  rename(Count = nuclear_weapons_stockpile) %>%
  mutate(
    x = coordinates[, 1],
    y = coordinates[, 2],
    pt_label = paste0(Entity, "\n", si_count(Count))) %>%
  # filter(Year == 1987) %>%
  split(.$Year) %>%
  iwalk(warheads_map)
```

We can manually stitch this together in a video with `ffmpeg`:

```{r videoout}
system2("ffmpeg", c(
  "-framerate", "3",
  "-pattern_type", "glob", "-i", "'warheads-*.png'",
  "-c:v", "libx264", "-pix_fmt", "yuv420p",
  "warheads.mp4"))
```